---
title: "data_processing_2"
author: "Charis Chan, Joyce Ching, and Inderpal Kaur"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
    
## Loading and processing raw data

```{r}
library(readxl)
library(readr)
library(data.table)
library(plyr)
library(dplyr)
library(lubridate)
library(stringi)
```

### Create new directories to hold the data created or used
```{r}
new_folders <- c('raw', 'external', 'interim','processed')
path <- "../../data"
for (folder in new_folders) {
  # check if the folder exists, if not then create folder
  ifelse(!dir.exists(file.path(path, folder)), dir.create(file.path(path, folder)), FALSE)
}

```

### Loading policy data at state level
```{r}
# download the state policy data and save to folder: /data/raw
state_policy_url <- "https://mids-w203.s3-us-west-1.amazonaws.com/COVID-19+US+state+policy+database+(CUSP).xlsx"
state_policy_destfile <- "../../data/raw/COVID_19_US_state_policy_database_CUSP.xlsx"
curl::curl_download(state_policy_url, state_policy_destfile)
```

```{r}
# read Excel sheet for Stay at Home data
names <- c("stateName", "stateCode", "stateFIPS", "order_start_date", "religious_exempt", "order_without_restriction_date", "order_end_date", "order_start_desc", "religious_exempt_desc", "order_without_restriction_desc", "order_end_desc")
types <- c("text", "text", "numeric", "date", "numeric", "date", "date", "skip", "text", "text", "text", "text")
stay_at_home <- read_xlsx(state_policy_destfile, sheet="Stay at Home", range="A2:L52", col_names=names, col_types=types)
stay_at_home <- stay_at_home[,1:7]

# convert dates into Date datatype
stay_at_home[] <- lapply(stay_at_home, function(x) if(is.POSIXct(x)) as.Date(x) else x)

# replace invalid dates with NAs
stay_at_home <- data.frame(sapply(stay_at_home, function(x) {gsub(as.Date("1899-12-31"), NA, x)}))

# fix datatypes
stay_at_home[,c("stateName", "stateCode")] <- apply(stay_at_home[,c("stateName", "stateCode")], 2, function(x) as.character(x))
stay_at_home[,c("stateFIPS","religious_exempt")] <- apply(stay_at_home[,c("stateFIPS","religious_exempt")], 2, function(x) as.numeric(as.character(x)))
stay_at_home[,c("order_start_date", "order_without_restriction_date", "order_end_date")] <- lapply(stay_at_home[,c("order_start_date", "order_without_restriction_date", "order_end_date")], as.Date)

head(stay_at_home)

##################

# read Excel sheet for Physical Distance Closures data
names <- c("stateName", "stateCode", "stateFIPS", "date_closed_schools", "date_closed_daycares", "date_banned_nursing_home_visitors", "date_closed_other_nonessential_businesses", "date_closed_restaurants", "date_closed_gyms", "date_closed_movie_theaters", "date_closed_bars")
types <- c("text", "text", "numeric", "date", "date", "date", "date", "date", "date", "date", "date")
physical_distance_closures <- read_xlsx(state_policy_destfile, sheet="Physical Distance Closures", range="A2:K52", col_names=names, col_types=types)

# convert dates into Date datatype
physical_distance_closures[] <- lapply(physical_distance_closures, function(x) if(is.POSIXct(x)) as.Date(x) else x)

# replace invalid dates with NAs
physical_distance_closures <- data.frame(lapply(physical_distance_closures, function(x) {gsub(as.Date("1899-12-31"), NA, x)}))

# fix datatypes
physical_distance_closures[,c("stateName", "stateCode")] <- apply(physical_distance_closures[,c("stateName", "stateCode")], 2, function(x) as.character(x))
physical_distance_closures[,c("stateFIPS")] <- as.numeric(as.character(physical_distance_closures[,c("stateFIPS")]))
physical_distance_closures[,c("date_closed_schools", "date_closed_daycares", "date_banned_nursing_home_visitors", "date_closed_other_nonessential_businesses", "date_closed_restaurants", "date_closed_gyms", "date_closed_movie_theaters", "date_closed_bars")] <- lapply(physical_distance_closures[,c("date_closed_schools", "date_closed_daycares", "date_banned_nursing_home_visitors", "date_closed_other_nonessential_businesses", "date_closed_restaurants", "date_closed_gyms", "date_closed_movie_theaters", "date_closed_bars")], as.Date)

head(physical_distance_closures)

#################################

# read Excel sheet for Reopening data
names <- (c("stateName", "stateCode", "stateFIPS", "reopen_businesses", "reopen_restaurants", "reopen_restaurants_outdoor_dining", "reopen_gyms", "reopen_movie_theaters", "reopen_hair_salons", "reopen_religious_gatherings", "reopen_other_nonessential", "reopen_bars", "reopen_childcare"))
types <- (c("text", "text", "numeric", "date", "date", "date", "date", "date", "date", "date", "date", "date", "date"))
reopening <- read_xlsx(state_policy_destfile, sheet="Reopening", range="A2:M52", col_names=names, col_types=types)

# convert dates into Date datatype
reopening[] <- lapply(reopening, function(x) if(is.POSIXct(x)) as.Date(x) else x)

# replace invalid dates with NAs
reopening <- data.frame(lapply(reopening, function(x) {gsub(as.Date("1899-12-31"), NA, x)}))

# fix datatypes
reopening[,c("stateName", "stateCode")] <- apply(reopening[,c("stateName", "stateCode")], 2, function(x) as.character(x))
reopening[,c("stateFIPS")] <- as.numeric(as.character(reopening[,c("stateFIPS")]))
reopening[,c("reopen_businesses", "reopen_restaurants", "reopen_restaurants_outdoor_dining", "reopen_gyms", "reopen_movie_theaters", "reopen_hair_salons", "reopen_religious_gatherings", "reopen_other_nonessential", "reopen_bars", "reopen_childcare")] <- lapply(reopening[,c("reopen_businesses", "reopen_restaurants", "reopen_restaurants_outdoor_dining", "reopen_gyms", "reopen_movie_theaters", "reopen_hair_salons", "reopen_religious_gatherings", "reopen_other_nonessential", "reopen_bars", "reopen_childcare")], as.Date)

head(reopening)

###############

# read Excel sheet for Face Masks data
names <- (c("stateName", "stateCode", "stateFIPS", "mask_use_by_all", "enforced_by_fines", "enforced_by_criminal_charge", "no_enforcement", "mask_use_by_employees", "end_mask_use_by_all", "state_gov_preventing_local"))
types <- (c("text", "text", "numeric", "date", "numeric", "numeric", "numeric", "date", "date", "date"))
face_masks <- read_xlsx(state_policy_destfile, sheet="Face Masks", range="A2:J52", col_names=names, col_types=types)

# convert dates into Date datatype
face_masks[] <- lapply(face_masks, function(x) if(is.POSIXct(x)) as.Date(x) else x)

# replace invalid dates with NAs
face_masks <- data.frame(lapply(face_masks, function(x) {gsub(as.Date("1899-12-31"), NA, x)}))

# fix datatypes
face_masks[,c("stateName", "stateCode")] <- apply(face_masks[,c("stateName", "stateCode")], 2, function(x) as.character(x))
face_masks[,c("stateFIPS", "enforced_by_fines", "enforced_by_criminal_charge", "no_enforcement")] <- apply(face_masks[,c("stateFIPS", "enforced_by_fines", "enforced_by_criminal_charge", "no_enforcement")], 2, function(x) as.numeric(as.character(x)))
face_masks[,c("mask_use_by_all", "mask_use_by_employees", "end_mask_use_by_all", "state_gov_preventing_local")] <- lapply(face_masks[,c("mask_use_by_all", "mask_use_by_employees", "end_mask_use_by_all", "state_gov_preventing_local")], as.Date)

head(face_masks)
```

### Loading data about state demographics and characteristics
```{r}
# download Excel file for state-level characteristics to folder: /data/raw
covid_state_url <- "https://mids-w203.s3-us-west-1.amazonaws.com/covid-19.xlsx"
covid_state_destfile <- "../../data/raw/covid-19.xlsx"
curl::curl_download(covid_state_url, covid_state_destfile)

# read Excel file
covid_19 <- read_xlsx(covid_state_destfile, sheet=2, skip=1)

# select important features
state_info_columns <- c('State', 'Governor', "Population density per square miles","Population 2018", 
                        "Nonelderly Adults Who Have A Pre-Existing Condition",
                        "Percent at risk for serious illness due to COVID",
                        "All-cause deaths 2018", "Number Homeless (2019)", 
                        "Medicaid Expenditures as a Percent of Total State Expenditures by Fund", 
                        "Life Expectancy at Birth (years)", "Percent Unemployed (2018)", 
                        "Percent living under the federal poverty line (2018)",
                        "Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars)",
                        "Median Annual Household Income", "Children 0-18", "Adults 19-25","Adults 26-34",
                        "Adults 35-54", "Adults 55-64", "65+")
state_info <- covid_19[1:51,state_info_columns]
colnames(state_info)[1] <- 'stateName'

# extract political party of state governors
state_info['governor_political_party'] <- stri_extract(stri_extract(state_info$Governor, regex = '\\(.\\)'), regex = '\\w')
state_info$governor_political_party <- as.factor(state_info$governor_political_party)

head(state_info)
```

### Joining state policy data for stay at home orders, closures, reopenings, face masks, and state info
```{r}
stayathome_physicaldistanceclosures <- join(stay_at_home, physical_distance_closures, by=c("stateName", "stateCode", "stateFIPS"), type="full")

stayathome_physicaldistanceclosures_reopening <- join(stayathome_physicaldistanceclosures, reopening, by=c("stateName", "stateCode", "stateFIPS"), type="full")

state_pol <- join(stayathome_physicaldistanceclosures_reopening, face_masks, by=c("stateName", "stateCode", "stateFIPS"), type="full")

state_policies <- join(state_pol, state_info, by = c('stateName'))
state_policies
```


### Loading mobility data at state level
```{r}
# download, unzip, and save the US mobility data to folder: /data/raw
mobility_url <- "https://www.gstatic.com/covid19/mobility/Region_Mobility_Report_CSVs.zip"
mobility_destfile <- "../../data/raw/Region_Mobility_Report_CSVs.zip"
us_mobility_destfile <- "2020_US_Region_Mobility_Report.csv"
curl::curl_download(mobility_url, mobility_destfile)
master <- as.character(unzip(mobility_destfile, list = TRUE)$Name)

# read the CSV file for US mobility data
mobility <- read.csv(unz(mobility_destfile, us_mobility_destfile), header = TRUE, sep = ",")

# convert dates into Date datatype
mobility$date <- as.Date(mobility$date)

# rename columns
colnames(mobility)[1:14] <- c("countryCode", "countryName", "stateName", "countyName", "metro_area", "iso", "countyFIPS", "date", "retail_and_recreation", "grocery_and_pharmacy", "parks", "transit_stations", "workplaces", "residential")
mobility <- mobility[, -c(2, 5)]
head(mobility, n=10)
```

```{r}
# select state-level mobility data
state_mobility_columns <- c("stateName", "date", "retail_and_recreation", "grocery_and_pharmacy", "parks",
                            "transit_stations","workplaces", "residential")
state_mobility <- mobility[mobility$iso != "", state_mobility_columns]

# rename columns
colnames(state_mobility)[3:8] <- c("state_retail_and_recreation", "state_grocery_and_pharmacy", "state_parks", "state_transit_stations", "state_workplaces", "state_residential")
```

### Loading state-level data on Covid-19 cases and deaths
```{r}
# download CSV for Covid-19 data to folder: /data/external
covid_states_url <- "https://drive.google.com/uc?export=download&id=1lm-JXbJh-f1JgmkeBjNTMv1I6jE-oLqW" 
covid_states_destfile <- "../../data/external/United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv"
curl::curl_download(covid_states_url, covid_states_destfile)

# read CSV
covid_states <- read_csv("./United_States_COVID-19_Cases_and_Deaths_by_State_over_Time.csv", col_types = cols(
  submission_date = 'c',
  state = 'c',
  tot_cases = 'i',
  conf_cases = 'i',
  prob_cases = 'i',
  new_case = 'i',
  pnew_case = 'i',
  tot_death = 'i',
  conf_death = 'i',
  prob_death = 'i',
  new_death = 'i',
  pnew_death = 'i',
  created_at = 'c',
  consent_cases = 'c',
  consent_deaths = 'c'
))

# convert dates to Date datatype
covid_states$submission_date <- as.Date(covid_states$submission_date, tryFormats = c("%m/%d/%y"))
head(covid_states)
```



### Joining information about Covid-19, mobility, and policies per day for each state
```{r}
# taking out some columns that will not be used for combining dataframes or in answering our research question
order_stay <- state_policies[,-which(names(state_policies) %in% c('stateFIPS','order_without_restriction_date', 'Governor'))]
colnames(order_stay)[1] <-  'stateName'

colnames(covid_states)[c(1,2)] <- c('date', 'stateCode')
# joining the covid cases/deaths info with order dates
covid_state_name <- join(covid_states, order_stay[c(1,2)], type = 'full', by = 'stateCode')

# joining state mobility with covid cases/death info and order dates
covid_mobility_cases <- join(state_mobility, covid_state_name, type = 'left', by = c('date', 'stateName'))
covid_country <- join(covid_mobility_cases, order_stay[-c(2)], type = 'left', by = 'stateName')
#take out all the states that never had a stay at home order
covid_country <- covid_country[!is.na(covid_country$order_start_date),]

```

### adding regions for states
```{r} 
# putting states in regions - from Census Bureau
# West =  Arizona, Colorado, Idaho, Montana, Nevada, New Mexico, Utah, Wyoming, Alaska, California, Hawaii, Oregon, and Washington
# Midwest = Illinois, Indiana, Michigan, Ohio, Wisconsin,Iowa, Kansas, Minnesota, Missouri, Nebraska, North Dakota, and South Dakota
# South = Delaware, Florida, Georgia, Maryland, North Carolina, South Carolina, Virginia, District of Columbia, West Virginia, Alabama, Kentucky, Mississippi, Tennessee, Arkansas, Louisiana, Oklahoma, and Texas
# Northeast = Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont, New Jersey, New York, and Pennsylvania


# initializing us_region
covid_country['US_Region'] <- ''
# west region
covid_country[covid_country$stateName %in% c('Arizona', 'Colorado', 'Idaho', 'Montana', 'Nevada', 'New Mexico', 'Utah', 'Wyoming', 'Alaska', 'California', 'Hawaii', 'Oregon', 'Washington'),'US_Region'] <- 'West'
# midwest
covid_country[covid_country$stateName %in% c('Illinois', 'Indiana', 'Michigan', 'Ohio', 'Wisconsin','Iowa', 'Kansas', 'Minnesota', 'Missouri', 'Nebraska', 'North Dakota', 'South Dakota'),'US_Region'] <- 'Midwest'
# south
covid_country[covid_country$stateName %in% c('Delaware', 'Florida', 'Georgia', 'Maryland', 'North Carolina', 'South Carolina', 'Virginia', 'District of Columbia', 'West Virginia', 'Alabama', 'Kentucky', 'Mississippi', 'Tennessee', 'Arkansas', 'Louisiana', 'Oklahoma', 'Texas'),'US_Region'] <- 'South'
# northeast
covid_country[covid_country$stateName %in% c('Connecticut', 'Maine', 'Massachusetts', 'New Hampshire', 'Rhode Island', 'Vermont', 'New Jersey', 'New York', 'Pennsylvania'),'US_Region'] <- 'Northeast'


# take out CA and NM because they have not ended the stay at home order
covid_country <- covid_country[!covid_country$stateName %in% c('California','New Mexico'),]
```


### Create function to calculate days since order
```{r}
days_since <- function(dates, order_start_date, order_end_date) {
  
  days_since_start <- dates - (order_start_date - 1)
  # order hasn't started yet - taking out the negatives
  days_since_start[days_since_start <0] <- 0
  
  days_since_end <- dates - order_end_date
  
  # order still hasn't ended - taking out the negatives
  days_since_end[days_since_end < 0] <- 0
  
  # order not in effect - marking the end of order
  days_since_start[days_since_end > 0] <- NA
  return(days_since_start)
}
```

### Use function above (days_since) to get the dates since orders started
```{r}
# Stay at home order
covid_country['order_length'] <- days_since(covid_country$date, covid_country$order_start_date, covid_country$order_end_date)


# daycare
covid_country['daycare_closed'] <- days_since(covid_country$date, covid_country$date_closed_daycares, covid_country$reopen_childcare)


# gyms
covid_country['gyms_closed'] <- days_since(covid_country$date, covid_country$date_closed_gyms, covid_country$reopen_gyms)


# restaurants
covid_country['restaurants_closed'] <- days_since(covid_country$date, covid_country$date_closed_restaurants, covid_country$reopen_restaurants)


# movie theaters
covid_country['movie_theaters_closed'] <- days_since(covid_country$date, covid_country$date_closed_movie_theaters, covid_country$reopen_movie_theaters)


# bars
covid_country['bars_closed'] <- days_since(covid_country$date, covid_country$date_closed_bars, covid_country$reopen_bars)



# other non essentials
covid_country['other_non_essentials_closed'] <- days_since(covid_country$date, covid_country$date_closed_other_nonessential_businesses, covid_country$reopen_other_nonessential)


# mask use by all
covid_country['mask_mandate'] <- days_since(covid_country$date, covid_country$mask_use_by_all, covid_country$end_mask_use_by_all)

# add covid country to the interim folder
write.csv(covid_country, "../../data/interim/covid_states_daily.csv")

```


### Comparing the mobility from the start and end of SIP
```{r} 
# columns we want for state information
columns_names <- c("stateName", 'US_Region', "Population density per square miles","Population 2018",
                   "Nonelderly Adults Who Have A Pre-Existing Condition",
                   "Percent at risk for serious illness due to COVID", "All-cause deaths 2018", 
                   "Number Homeless (2019)", 
                   "Medicaid Expenditures as a Percent of Total State Expenditures by Fund", 
                   "Life Expectancy at Birth (years)", "Percent Unemployed (2018)", 
                   "Percent living under the federal poverty line (2018)",
                   "Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars)",
                   "Median Annual Household Income", "Children 0-18", "Adults 19-25","Adults 26-34",
                   "Adults 35-54", "Adults 55-64", "65+", "governor_political_party", "order_length",
                   "daycare_closed","gyms_closed", "restaurants_closed", "movie_theaters_closed", 
                   "bars_closed", "other_non_essentials_closed","mask_mandate",'order_start_date')

mobility_change <- covid_country[(covid_country$date == covid_country$order_end_date) , columns_names]


# getting only the columns about mobility to subtract start with the end mobility
mobility_columns <- c("state_retail_and_recreation","state_grocery_and_pharmacy",
                      "state_parks","state_transit_stations","state_workplaces","state_residential")

columns_state_mobility <- c('stateName', mobility_columns)

# get the mobility rates for the week of when the order started and ended
start_state_mobility <- covid_country[(covid_country$date >=covid_country$order_start_date) & (covid_country$date < covid_country$order_start_date+7),columns_state_mobility] 

end_state_mobility <- covid_country[(covid_country$date >covid_country$order_end_date-7) & (covid_country$date <= covid_country$order_end_date),columns_state_mobility] 

# average the week's mobility rate
avg_start_mobility <- start_state_mobility %>%
  group_by(stateName) %>%
  summarise_all(mean)

avg_end_mobility <- end_state_mobility %>%
  group_by(stateName) %>%
  summarise_all(mean)
new_names_for_mobility <- c("retail_and_recreation_diff","grocery_and_pharmacy_diff",
                      "parks_diff","transit_stations_diff","workplaces_diff","residential_diff")

# subtract the end mobility rate from the start mobility
mobility_change[, new_names_for_mobility]<- avg_start_mobility[,mobility_columns] -avg_end_mobility[,mobility_columns]


# find the order in which states first 
mobility_change['order_start_rank'] <- min_rank(mobility_change$order_start_date)

# drop order_start_date since we have the order_start_rank
mobility_change <- mobility_change[, !(colnames(mobility_change) %in% c('order_start_date'))]
```

```{r}
mobility_change
```


### getting average new cases/deaths per day for that time period, need population
```{r}
covid_avg <-  covid_country %>%
  filter(order_length > 0) %>% 
  group_by(stateName) %>%
  dplyr::summarise(avg_new_cases = mean(new_case, na.rm = TRUE), 
            avg_new_deaths = mean(new_death, na.rm = TRUE), 
            avg_new_cases_per_100K = mean(new_case/(`Population 2018`/100000),na.rm = TRUE), 
            new_deaths_per_100K = mean(new_death/(`Population 2018`/100000),na.rm = TRUE), 
            max_new_cases = max(new_case, na.rm = TRUE), 
            max_new_deaths = max(new_death, na.rm = TRUE))




covid_avg_states <- join(mobility_change, covid_avg, by = 'stateName')


covid_avg_states
```

```{r}
# save the covid state data into the processed folder
write.csv(covid_avg_states, "../../data/processed/covid_states.csv")
```

